Simulations of Ballot Polling Risk-Limiting Audits

### 2022-05-16 providence/sections/appendix.tex
prog

\begin{definition}[$(\alpha,p_a, p_0,k_{j-1},n_{j-1},n_j)$-\Providence]
    \label{def:minervatwo}
    For cumulative round size $n_i$ for round $i$ and a cumulative $k_i$ ballots for the reported winner found in round $i$, the \R \Providence stopping rule for the $j^{th}$ round is:
$$
\mathcal{A}(X_{j})=  \left\{ \begin{array}{ll} \text{Correct} ~~~~ \omega_{j}(k_{j}, k_{j-1}, p_a, p_0, n_j, n_{j-1}) \geq \frac{1}{\alpha}\\
        % & \\
        % incorrect& ~~~ \sigma_n < \frac{\beta}{1-\alpha} \\
        % & \\
        Undetermined ~~else \\
    \end{array}
    \right .
$$
where $\omega _{1}\triangleq \tau_{1}$ and for $j\ge 2$, we define $\omega _{j}$ as follows:
\begin{equation}
    \begin{aligned}
    \omega_{j}(k_{j}, k_{j-1}, p_a, p_0, n_{j}, n_{j-1})\triangleq
    \sigma(k_{j-1},p_a,p_0,n_{j-1})\cdot \tau_1(k_{j}-k_{j-1},p_a,p_0,n_j-n_{j-1})
    \end{aligned}
\end{equation}
\end{definition}
% old definition of omega_j for j\ge 2

Notice that for $j\ge 2$, unlike $\tau_j$, computing $\omega_j$ requires no convolution.

\section{Proofs: Minerva 2.0 is Risk Limiting}
\begin{lemma}
    \label{lemma:sigma_increasing}
For $0<p_0< p_a< 1$ and $n>0$, the ratio $\sigma(k,p_a,p_0,n)$ is strictly increasing as a function of $k$ for $0\le k\le n$.
\end{lemma}
\begin{proof}
From $0<p_0< p_a< 1$, we get
$$
\frac{p_a}{p_0}>1
$$
and
$$
1-p_0>1-p_a\implies \frac{1-p_0}{1-p_a}>1,
$$
and thus
$$
\frac{p_a(1-p_0)}{p_0(1-p_a)}>1.
$$
Now simply observe that
$$
\frac{p_a(1-p_0)}{p_0(1-p_a)}\cdot\sigma(k, p_a, p_0, n)
=\frac{p_a(1-p_0)}{p_0(1-p_a)}\cdot\frac{p_a^{k} (1-p_a)^{n-k}}{p_a^{k} (1-p_a)^{n-k}}
=\frac{p_a^{k+1} (1-p_a)^{n-(k+1)}}{p_a^{k+1} (1-p_a)^{n-(k+1)}}
=\sigma(k+1, p_a, p_0, n).
$$
\end{proof}

@@

%Here we introduce the notion of the history of an audit.
%In round $j+1$, all previous round sizes and cumulative
%tallies of winner ballots are together relevant information to the new round.
%We call this information the history of the audit through round $j$,
%and we denote it as follows:
%$$\bm{k}_j\triangleq(k_1,k_2,\ldots,k_j,n_1,n_2,\ldots,n_j)$$
\begin{lemma}
\label{lemma:any_ratio_is_sigma_simple}
For $j\ge 1$,
$$\frac{Pr[\bm{K_j}=\bm{k_j} \mid \bm{n_j}, H_a]}{Pr[\bm{K_j}=\bm{k_j} \mid \bm{n_j}, H_0]} = \sigma(k_j, p_a, p_0, n_j).$$
%$$\frac{Pr[K_j = k_{j} \mid n_{j-1}^*, k_{j-1}^*, H_a]}{Pr[K_j = k_{j} \mid n_{j-1}^*, k_{j-1}^*, H_0]} = \sigma(k_j, p_a, p_0, n_j).$$
\end{lemma}
\begin{proof}
We induct on the number of rounds.
For $j=1$, we have
$$\frac{Pr[\bm{K_1}=\bm{k_1} \mid \bm{n_1},H_a]}{Pr[\bm{K_1}=\bm{k_1} \mid  \bm{n_1},H_0]} =\frac{Pr[K_1 = k_{1} \mid n_1,H_a]}{Pr[K_1 = k_1 \mid n_1,H_0]} = \frac{\text{Bin}(k_1,n_1,p_a)}{\text{Bin}(k_1,n_1,p_0)}=\sigma(k_1, p_a, p_0, n_1).$$
%
% Suppose the lemma is true for $j=m$.
% That is,
% $$\frac{Pr[\bm{K_m}=\bm{k_m} \mid H_a]}{Pr[\bm{K_m}=\bm{k_m} \mid  H_0]} = \sigma(k_m, p_a, p_0, n_m).$$
%
Suppose the lemma is true for round $j=m$ with history $\bm{k_m}$.
%We will show it is true for $j=m+1$.
Observe that
 $$\frac{Pr[\bm{K_{m+1}}=\bm{k_{m+1}} \mid \bm{n_{m+1}},H_a]}{Pr[\bm{K_{m+1}}=\bm{k_{m+1}} \mid \bm{n_{m+1}}, H_0]} = \frac{ Pr[\bm{K_{m}}=\bm{k_{m}}\mid \bm{n_{m+1}},H_a] \cdot Pr[K_{m+1}'=k_{m+1}'|\bm{k_m},\bm{n_{m+1}},H_a]}{ Pr[\bm{K_{m}}= \bm{k_{m}} \mid  \bm{n_{m+1}},H_0]  \cdot  Pr[K_{m+1}'=k_{m+1}'|\bm{k_m},\bm{n_{m+1}},H_0]  }$$
 $$=\sigma(k_m, p_a, p_0, n_m) \cdot \frac{Pr[K_{m+1}'=k_{m+1}'|\bm{k_m}, \bm{n_{m+1}}, H_a]}{Pr[K_{m+1}'=k_{m+1}'|\bm{k_m},\bm{n_{m+1}},H_0]}$$
 by the induction hypothesis.
Then this is simply equal to
 $$\sigma(k_m, p_a, p_0, n_m)\cdot\frac{\text{Bin}(k_{m+1}',n_{m+1}',p_a)}{\text{Bin}(k_{m+1}',n_{m+1}',p_0)}
 $$$$
 =\frac{p_a^{k_m} (1-p_a)^{n_m-k_m}}{p_0^{k_m} (1-p_0)^{n_m-k_m}} \cdot
 \frac{p_a^{k_{m+1}'} (1-p_a)^{n_{m+1}'-k_{m+1}'}}{p_0^{k_{m+1}'} (1-p_0)^{n_{m+1}'-k_{m+1}'}}
 $$
 $$
 =\sigma(k_{m+1}, p_a, p_0, n_{m+1})
 $$
\end{proof}

\begin{theorem}
\label{thm:minerva2_is_rla_new}
An $(\alpha,p_a, p_0,k_{j-1},n_{j-1},n_j)$-\Providence audit is an
$\alpha$-RLA.
\end{theorem}
\begin{proof}
Let $\mathcal{A}=(\alpha,p_a, p_0,k_{j-1},n_{j-1},n_j)$-\Providence.
Let $\bm{n_j}$ be the cumulative roundsizes used in this
audit, with corresponding cumulative tallies of
ballots for the reported winner $\bm{k_j}$.
For round $j=1$, by Definitions \ref{def:minervatwo}
and \ref{def:minerva_ratio}, we see that
the $\mathcal{A}=\text{Correct}$ (the audit stops) only when
$$
\tau_1(k_{1},p_a,p_0,n_1)
=\frac{Pr[K_{1} \geq k_{1} \mid H_a, n_1]}{Pr[K_{1} \geq k_{1} \mid H_0, n_1]}
\ge \frac{1}{\alpha}.
$$
By Lemma \ref{lemma:minerva2_kmin_exists}, we see that this
is equivalent to the following:
$$
\frac{Pr[K_{1} \geq k_{min,1} \mid H_a, n_1]}{Pr[K_{1} \geq k_{min, 1} \mid H_0, n_1]}
\ge \frac{1}{\alpha}.
$$

For any round $j\ge 2$, by Definition \ref{def:minervatwo}
and Lemma \ref{lemma:minerva2_kmin_exists},
$\mathcal{A}=\text{Correct}$ (the audit stops) only when
$$
\omega_{j}(k_{j}, k_{j-1}, p_a, p_0, n_{j}, n_{j-1}, \alpha )\triangleq
\sigma(k_{j-1},p_a,p_0,n_{j-1})\cdot \tau_1(k_{j}-k_{j-1},p_a,p_0,n_j-n_{j-1})
\ge \frac{1}{\alpha}.
$$
%Let $d_j=k_j-k_{j-1}$ be shorthand for the new draw.
%could do this for simpler notation^
By Lemma \ref{lemma:any_ratio_is_sigma_simple}
and Definition \ref{def:minerva_ratio}, this is equivalent to
$$
\frac{\Pr[\bm{k_{j-1}} \mid H_a]\cdot Pr[K_{j} \ge k_{j} \mid \bm{k_{j-1}}, H_a, n{j}]}{\Pr[\bm{k_{j-1}} \mid H_0]\cdot Pr[K_{j} \ge k_{j} \mid \bm{k_{j-1}}, H_0, n{j}]}\ge \frac{1}{\alpha}.
$$
By Lemma~\ref{lemma:minerva2_kmin_exists} and Definition~\ref{def:minervatwo},
we see that there exists a $k_{min, j}\le k_j$ for which
$$
\frac{\Pr[\bm{k_{j-1}} \mid H_a]\cdot Pr[K_{j} \ge k_{min, j} \mid \bm{k_{j-1}}, H_a, n{j}]}{\Pr[\bm{k_{j-1}} \mid H_0]\cdot Pr[K_{j} \ge k_{min, j} \mid \bm{k_{j-1}}, H_0, n{j}]} \ge
\frac{\Pr[\bm{k_{j-1}} \mid H_a]\cdot Pr[K_{j} \ge k_{j} \mid \bm{k_{j-1}}, H_a, n{j}]}{\Pr[\bm{k_{j-1}} \mid H_0]\cdot Pr[K_{j} \ge k_{j} \mid \bm{k_{j-1}}, H_0, n{j}]}\ge \frac{1}{\alpha}.
$$

%Equivalently,
%$$
%\frac{Pr[K_{j} \ge k_{min, j}\wedge K_{j-1} = k_{j-1} \mid n_{j}^*, k_{j-2}^*, H_a]}{Pr[K_{j}\ge k_{min, j} \wedge K_{j-1} = k_{j-1} \mid n_{j}^*, k_{j-2}^*, H_0]}\ge \frac{1}{\alpha}.
%$$
%Taking the sum over all possible rounds and corresponding preceding values
%of $k$, we get
Taking the sum over all possible audit histories, we get
$$
\frac{\sum_{\bm{k_j}}\Pr[\bm{k_{j-1}} \mid H_a]\cdot Pr[K_{j} \ge k_{min, j} \mid \bm{k_{j-1}}, H_a, n{j}]}{\sum_{\bm{k_j}}\Pr[\bm{k_{j-1}} \mid H_0]\cdot Pr[K_{j} \ge k_{min, j} \mid \bm{k_{j-1}}, H_0, n{j}]}\ge \frac{1}{\alpha}.
$$
Finally, because the total probability of stopping the audit under
the alternative hypothesis is less than 1, we get
$$
\frac{Pr[\mathcal{A}=\text{Correct} \mid H_a]}{Pr[\mathcal{A}=\text{Correct} \mid H_0]}\ge \frac{1}{\alpha}
$$
$$
Pr[\mathcal{A}=\text{Correct} \mid H_0]
\le
Pr[\mathcal{A}=\text{Correct} \mid H_a] \cdot \alpha
\le
\alpha.
$$
\end{proof}

\subsection{Resistance against an adversary choosing round sizes}

>>>>

\begin{lemma}
    \label{lemma:sigma_increasing}
For $0<p_0< p_a< 1$ and $n>0$, the ratio $\sigma(k,p_a,p_0,n)$ is strictly increasing as a function of $k$ for $0\le k\le n$.
\end{lemma}
\begin{proof}
From $0<p_0< p_a< 1$, we get
$$
\frac{p_a}{p_0}>1
$$
and
$$
1-p_0>1-p_a\implies \frac{1-p_0}{1-p_a}>1,
$$
and thus
$$
\frac{p_a(1-p_0)}{p_0(1-p_a)}>1.
$$
Now simply observe that
$$
\frac{p_a(1-p_0)}{p_0(1-p_a)}\cdot\sigma(k, p_a, p_0, n)
=\frac{p_a(1-p_0)}{p_0(1-p_a)}\cdot\frac{p_a^{k} (1-p_a)^{n-k}}{p_a^{k} (1-p_a)^{n-k}}
=\frac{p_a^{k+1} (1-p_a)^{n-(k+1)}}{p_a^{k+1} (1-p_a)^{n-(k+1)}}
=\sigma(k+1, p_a, p_0, n).
$$
\end{proof}

@@

\begin{lemma}
\label{lemma:any_ratio_is_sigma_simple}
For $j\ge 1$,
$$\frac{Pr[\bm{K_j}=\bm{k_j} \mid \bm{n_j}, H_a]}{Pr[\bm{K_j}=\bm{k_j} \mid \bm{n_j}, H_0]} = \sigma(k_j, p_a, p_0, n_j).$$
\end{lemma}
\begin{proof}
We induct on the number of rounds.
For $j=1$, we have
$$\frac{Pr[\bm{K_1}=\bm{k_1} \mid \bm{n_1},H_a]}{Pr[\bm{K_1}=\bm{k_1} \mid  \bm{n_1},H_0]} =\frac{Pr[K_1 = k_{1} \mid n_1,H_a]}{Pr[K_1 = k_1 \mid n_1,H_0]} = \frac{\text{Bin}(k_1,n_1,p_a)}{\text{Bin}(k_1,n_1,p_0)}=\sigma(k_1, p_a, p_0, n_1).$$
Suppose the lemma is true for round $j=m$ with history $\bm{k_m}$.
Observe that
 $$\frac{Pr[\bm{K_{m+1}}=\bm{k_{m+1}} \mid \bm{n_{m+1}},H_a]}{Pr[\bm{K_{m+1}}=\bm{k_{m+1}} \mid \bm{n_{m+1}}, H_0]} = \frac{ Pr[\bm{K_{m}}=\bm{k_{m}}\mid \bm{n_{m+1}},H_a] \cdot Pr[K_{m+1}'=k_{m+1}'|\bm{k_m},\bm{n_{m+1}},H_a]}{ Pr[\bm{K_{m}}= \bm{k_{m}} \mid  \bm{n_{m+1}},H_0]  \cdot  Pr[K_{m+1}'=k_{m+1}'|\bm{k_m},\bm{n_{m+1}},H_0]  }$$
 $$=\sigma(k_m, p_a, p_0, n_m) \cdot \frac{Pr[K_{m+1}'=k_{m+1}'|\bm{k_m}, \bm{n_{m+1}}, H_a]}{Pr[K_{m+1}'=k_{m+1}'|\bm{k_m},\bm{n_{m+1}},H_0]}$$
 by the induction hypothesis.
Then this is simply equal to
 $$\sigma(k_m, p_a, p_0, n_m)\cdot\frac{\text{Bin}(k_{m+1}',n_{m+1}',p_a)}{\text{Bin}(k_{m+1}',n_{m+1}',p_0)}
 $$$$
 =\frac{p_a^{k_m} (1-p_a)^{n_m-k_m}}{p_0^{k_m} (1-p_0)^{n_m-k_m}} \cdot
 \frac{p_a^{k_{m+1}'} (1-p_a)^{n_{m+1}'-k_{m+1}'}}{p_0^{k_{m+1}'} (1-p_0)^{n_{m+1}'-k_{m+1}'}}
 $$
 $$
 =\sigma(k_{m+1}, p_a, p_0, n_{m+1})
 $$
\end{proof}

### 2022-05-16 providence/sections/intro.tex
prog

Both \BRAVO and \Minerva have been integrated into election audit software {\em Arlo} \cite{arlo}, and, as such, are available for use in real election audits. Both have been used in real election audits\cite{usenix_minerva, virginia_rla}. For this reason, it is important to understand their properties over multiple rounds.

\subsection{Our Contributions}
We present \Providence, and provide the following:
\begin{enumerate}
\item Proof that \Providence is an RLA and resistant to an adversary who can choose subsequent round sizes with knowledge of previous samples
\item Simulations of \Providence, \Minerva, SO \BRAVO, and EoR \BRAVO which show that \Providence has similar efficiency to \Minerva, both greater than either implementation of \BRAVO
\item Results and analysis from the use of \Providence in a pilot audit in Rhode Island
\item Open source implementation of \Providence including the novel metric Probability of Misleading(name?)
\end{enumerate}

\subsection{Organization} 
Section \ref{sec:related} describes related work. The experiments we performed are described in section \ref{sec:expts}, and sections \ref{sec:results1} and \ref{sec:results2} present our results. Section \ref{sec:conc} has our conclusions.

>>>>

While more efficient than \BRAVO, \Minerva requires that the round schedule is fixed in advance of the audit. This may lead to significant unnecessary work. For example, suppose the fixed round schedule for \Minerva is $10,000$ ballots per round (a reasonable number in a state-wide contest with a narrow margin), and the first round sample finds a number of ballots for the winner just short of that necessary for the audit to stop. It then may be sufficient to draw a very small second round and still stop with high probability in the second round, but the additional $10,000$ ballots must be drawn. In contrast, subsequent \BRAVO round sizes can be chosen based on preceding samples.

An open question is whether a ballot polling RLA exists with the efficiency of \Minerva and this flexibility of \BRAVO.

\subsection{Our Contributions}
We present \Providence, and provide the following:
\begin{enumerate}
\item Proof that \Providence is an RLA and resistant to an adversary who can choose subsequent round sizes with knowledge of previous samples
\item Simulations of \Providence, \Minerva, SO \BRAVO, and EoR \BRAVO which show that \Providence has similar efficiency to \Minerva, both greater than either implementation of \BRAVO
\item Results and analysis from the use of \Providence in a pilot audit in Rhode Island
\item Open source implementation of \Providence 
%including the novel metric Probability of Misleading(name?)
\end{enumerate}

%\subsection{Organization} 
%Section \ref{sec:related} describes related work. The experiments we performed are described in section \ref{sec:expts}, and sections \ref{sec:results1} and \ref{sec:results2} present our results. Section \ref{sec:conc} has our conclusions.

### 2022-05-16 providence/sections/prov.tex
prog

\section{Proofs: Minerva 2.0 is Risk Limiting}
\begin{lemma}
    \label{lemma:sigma_increasing}
For $0<p_0< p_a< 1$ and $n>0$, the ratio $\sigma(k,p_a,p_0,n)$ is strictly increasing as a function of $k$ for $0\le k\le n$.
\end{lemma}
\begin{proof}
From $0<p_0< p_a< 1$, we get
$$
\frac{p_a}{p_0}>1
$$
and
$$
1-p_0>1-p_a\implies \frac{1-p_0}{1-p_a}>1,
$$
and thus
$$
\frac{p_a(1-p_0)}{p_0(1-p_a)}>1.
$$
Now simply observe that
$$
\frac{p_a(1-p_0)}{p_0(1-p_a)}\cdot\sigma(k, p_a, p_0, n)
=\frac{p_a(1-p_0)}{p_0(1-p_a)}\cdot\frac{p_a^{k} (1-p_a)^{n-k}}{p_a^{k} (1-p_a)^{n-k}}
=\frac{p_a^{k+1} (1-p_a)^{n-(k+1)}}{p_a^{k+1} (1-p_a)^{n-(k+1)}}
=\sigma(k+1, p_a, p_0, n).
$$
\end{proof}

\begin{lemma}
    \label{lemma:frac_sums_increasing}
    Given a monotone increasing sequence: $\frac{a_1}{b_1}, \frac{a_2}{b_2}, \ldots, \frac{a_n}{b_n}$, for $a_i, b_i > 0$, the sequence:
    $$z_i = \frac{\sum_{j=i}^n a_j}{\sum_{j=i}^n b_j}$$
    is also monotone increasing.
\end{lemma}

\begin{proof}
    Note that $z_i$ is a weighted average of the values of $\frac{a_j}{b_j}$ for $j \geq i$:
    $$z_i = \sum_{j=i}^n y_j \frac{a_j}{b_j}$$  ~for~ $$y_j = \frac{b_j}{\sum _{j=i}^n b_j} > 0.$$
    Further, $\sum _{j=i}^n y_j = 1$ and hence  $y_j \leq 1$ ~and~ $y_j = 1 \Leftrightarrow i=j=n$.
    Observe that, because $\frac{a_i}{b_i}$ is monotone increasing,
    $z_i \geq \frac{a_i}{b_i}$
    with equality if and only if $i=n$.
    Suppose $i < n$. Then
    $$z_{i+1} \geq \frac{a_{i+1}}{b_{i+1}} > \frac{a_i}{b_i}, $$
    and
    $$z_i = y_i \frac{a_i}{b_i} + (1-y_i) z_{i+1} < z_{i+1}.$$
    Thus $z_i$ is also monotone increasing.
\end{proof}

\begin{lemma}
    \label{lemma:tau1_increasing}
For $0<p_0< p_a< 1$ and $n>0$, the ratio $\tau_1(k,p_a,p_0,n)$ is strictly increasing as a function $k$ for $0\le k\le n$.
\end{lemma}
\begin{proof}
    Apply Lemmas \ref{lemma:sigma_increasing}-\ref{lemma:frac_sums_increasing}.
\end{proof}
\begin{lemma}
    \label{lemma:imin-exists}
    Given a strictly monotone increasing sequence: $x_1, x_2, \ldots x_n $ and some constant $A$,
    %there exists an $i_{min}$ such that
    $$A \le x_i \Leftrightarrow \exists i_{min} \le i ~\text{s.t.}~   x_{i_{min} -1} < A \le x_{i_{min}} \le x_{i},$$
    unless $A\le x_1$, in which case $i_{min} =1 $.
\end{lemma}
\begin{proof}
    Evident.
\end{proof}

\begin{lemma}
    \label{lemma:minerva2_kmin_exists}
    For $\mathcal{A}=(\alpha,p_a, p_0,k_{j-1},n_{j-1},n_j)$-\Providence, there exists\\ a $k_{min,j}(\Providence, p_a, p_0, k_{j-1}, n_{j-1}, n_j)$ such that $$\mathcal{A}(X_j)=\text{Correct}\iff k_j\ge k_{min,j}(\Providence,  \bm{n_j}, p_a, p_0).$$
\end{lemma}
\begin{proof}
    From Definition~\ref{def:minervatwo}, $$\mathcal{A}(X_j)=\text{Correct}\iff \omega_j(k_{j}, k_{j-1}, p_a, p_0, n_j, n_{j-1}) \ge \frac{1}{\alpha}.$$
    Now to apply Lemma~\ref{lemma:imin-exists}, it suffices to show that
    $\omega_j$ is monotone increasing with respect to $k_j$.
    For $j=1$, we have $\omega_1=\tau_1$, so $\omega_1$ is strictly increasing by Lemma \ref{lemma:tau1_increasing}. For $j\ge 2$,
    $$\omega_j(k_j,k_{j-1},p_a,p_0,n_j,n_{j-1},\alpha)=\sigma(k_{j-1},p_a,p_0,n_{j-1})\cdot \tau_1(k_{j}-k_{j-1},p_a,p_0,n_j-n_{j-1}).$$
    As a function of $k_j$, $\sigma$ is constant, and thus $\omega$ is strictly increasing by Lemma \ref{lemma:tau1_increasing}. Therefore by Lemma \ref{lemma:imin-exists}, we have the desired property.
\end{proof}

%Here we introduce the notion of the history of an audit.
%In round $j+1$, all previous round sizes and cumulative
%tallies of winner ballots are together relevant information to the new round.
%We call this information the history of the audit through round $j$,
%and we denote it as follows:
%$$\bm{k}_j\triangleq(k_1,k_2,\ldots,k_j,n_1,n_2,\ldots,n_j)$$
\begin{lemma}
\label{lemma:any_ratio_is_sigma_simple}
For $j\ge 1$,
$$\frac{Pr[\bm{K_j}=\bm{k_j} \mid \bm{n_j}, H_a]}{Pr[\bm{K_j}=\bm{k_j} \mid \bm{n_j}, H_0]} = \sigma(k_j, p_a, p_0, n_j).$$
%$$\frac{Pr[K_j = k_{j} \mid n_{j-1}^*, k_{j-1}^*, H_a]}{Pr[K_j = k_{j} \mid n_{j-1}^*, k_{j-1}^*, H_0]} = \sigma(k_j, p_a, p_0, n_j).$$
\end{lemma}
\begin{proof}
We induct on the number of rounds.
For $j=1$, we have
$$\frac{Pr[\bm{K_1}=\bm{k_1} \mid \bm{n_1},H_a]}{Pr[\bm{K_1}=\bm{k_1} \mid  \bm{n_1},H_0]} =\frac{Pr[K_1 = k_{1} \mid n_1,H_a]}{Pr[K_1 = k_1 \mid n_1,H_0]} = \frac{\text{Bin}(k_1,n_1,p_a)}{\text{Bin}(k_1,n_1,p_0)}=\sigma(k_1, p_a, p_0, n_1).$$
%
% Suppose the lemma is true for $j=m$.
% That is,
% $$\frac{Pr[\bm{K_m}=\bm{k_m} \mid H_a]}{Pr[\bm{K_m}=\bm{k_m} \mid  H_0]} = \sigma(k_m, p_a, p_0, n_m).$$
%
Suppose the lemma is true for round $j=m$ with history $\bm{k_m}$.
%We will show it is true for $j=m+1$.
Observe that
 $$\frac{Pr[\bm{K_{m+1}}=\bm{k_{m+1}} \mid \bm{n_{m+1}},H_a]}{Pr[\bm{K_{m+1}}=\bm{k_{m+1}} \mid \bm{n_{m+1}}, H_0]} = \frac{ Pr[\bm{K_{m}}=\bm{k_{m}}\mid \bm{n_{m+1}},H_a] \cdot Pr[K_{m+1}'=k_{m+1}'|\bm{k_m},\bm{n_{m+1}},H_a]}{ Pr[\bm{K_{m}}= \bm{k_{m}} \mid  \bm{n_{m+1}},H_0]  \cdot  Pr[K_{m+1}'=k_{m+1}'|\bm{k_m},\bm{n_{m+1}},H_0]  }$$
 $$=\sigma(k_m, p_a, p_0, n_m) \cdot \frac{Pr[K_{m+1}'=k_{m+1}'|\bm{k_m}, \bm{n_{m+1}}, H_a]}{Pr[K_{m+1}'=k_{m+1}'|\bm{k_m},\bm{n_{m+1}},H_0]}$$
 by the induction hypothesis.
Then this is simply equal to
 $$\sigma(k_m, p_a, p_0, n_m)\cdot\frac{\text{Bin}(k_{m+1}',n_{m+1}',p_a)}{\text{Bin}(k_{m+1}',n_{m+1}',p_0)}
 $$$$
 =\frac{p_a^{k_m} (1-p_a)^{n_m-k_m}}{p_0^{k_m} (1-p_0)^{n_m-k_m}} \cdot
 \frac{p_a^{k_{m+1}'} (1-p_a)^{n_{m+1}'-k_{m+1}'}}{p_0^{k_{m+1}'} (1-p_0)^{n_{m+1}'-k_{m+1}'}}
 $$
 $$
 =\sigma(k_{m+1}, p_a, p_0, n_{m+1})
 $$
\end{proof}

>>>>

Now we prove that \Providence is risk-limiting.

### 2022-05-16 providence/sections/sims.tex
prog

% simulations
We use simulations to provide additional evidence for theoretical claims and provide insight into audit behavior. As in \cite{sims}, we use margins from the 2020 US Presidential election, state-wide pairwise margins between the leading two candidates of 5\% or more. Narrower margins are computationally expensive, especially for the simulations with an underyling tie which quickly increase in sample size. We use the simulator in the R2B2 software library\cite{r2b2}. We perform $10000=10^4$ trials per margin for both an underyling announced outcomes and an underyling tie.

@@

Simulations with the underlying ballot distribution as announced provide insight into stopping probability and number of ballots drawn. Figure~\ref{prov-sprob} shows that the stoping probabilities over the first rounds are near and slightly above $90\%$ as expected, since our software chose round sizes to give at least a $90\%$ conditional stopping probability.
Figure~\ref{prov-asn} shows that the probability of stopping as a function of number of ballots sampled. Points above (higher probability of stopping) and to the left (fewer ballots) represent more efficient audits. As shown, \Providence has comparable efficiency to \Minerva, while both are significantly more efficient than either implementation of \BRAVO. Note that a difference in twice as many ballots could be negligible in a contest with a wide margin. In a contest with a narrow margin (in the 2020 US Presidential election, several states had margins less than $3\%$) the difference in number of ballots sampled could be weeks of work. Section~\ref{sec:workload} discusses workload in more depth.
%TODO add a more specific number of states for small margin

>>>>

% simulations
We use simulations to provide additional evidence for theoretical claims and gain insight into audit behavior. As in \cite{sims}, we use margins from the 2020 US Presidential election, state-wide pairwise margins between the leading two candidates of 5\% or more. Narrower margins are computationally expensive, especially for the simulations with an underyling tie which quickly increase in sample size. We use the simulator in the R2B2 software library\cite{r2b2}. We perform $10000=10^4$ trials per margin for both an underyling announced outcomes and an underyling tie.

@@

Simulations with the underlying ballot distribution as announced provide insight into stopping probability and number of ballots drawn. Figure~\ref{fig:prov-sprob} shows that the stoping probabilities over the first rounds are near and slightly above $90\%$ as expected, since our software chose round sizes to give at least a $90\%$ conditional stopping probability.
Figure~\ref{fig:prov-asn} shows that the probability of stopping as a function of number of ballots sampled. Points above (higher probability of stopping) and to the left (fewer ballots) represent more efficient audits. As shown, \Providence has comparable efficiency to \Minerva, while both are significantly more efficient than either implementation of \BRAVO. Note that a difference in twice as many ballots could be negligible in a contest with a wide margin. In a contest with a narrow margin (in the 2020 US Presidential election, several states had margins less than $3\%$) the difference in number of ballots sampled could be weeks of work. Section~\ref{sec:workload} discusses workload in more depth.
%TODO add a more specific number of states for small margin
